---
title: "White Wine"
author: "Saleh Saeed"
date: "January 17, 2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
install.packages('corrplot')
install.packages("munsell")
install.packages('ggplot2')
library(ggthemes)
library(ggplot2)
library(corrplot)
library(reshape2)
library(grid)
library(gridExtra)
```



```{r echo=FALSE}
getwd()
list.files()
wine <- read.csv("wineQualityWhites.csv",stringsAsFactors = FALSE)


```


```{r echo=FALSE}
str(wine)

```

```{r echo=FALSE}
summary(wine)
```


# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}

plot.hist('volatile.acidity', wine, 0.01,'Volatile Acidity',"   Volatile Acidity Distribution")
```
Volatile acidity is right skewed with many outliers
```{r echo=FALSE}
plot.hist('citric.acid', wine, 0.01,'Citric Acid ',"           Citric Acid Distribution")
```
Citric acid is normally distributed with many outliers 
```{r echo=FALSE}

plot.hist('residual.sugar', wine, 0.3,'Residual Sugar',"       Residual Sugar Distribution")
```

Residual sugar is highly right skewed, with less outliers.

```{r echo=FALSE}
ggplot(data = wine, aes(x=wine$residual.sugar))  +geom_histogram() +
scale_x_log10()
```
The histogram shows that the residual sugar has two peaks at 2 and 8

```{r echo=FALSE}
plot.hist('chlorides', wine, 0.001,'Chlorides',"       Chlorides Distribution")
```
Chlorides is normally distributed with many outliers

```{r echo=FALSE}
qplot(x = free.sulfur.dioxide, data = wine)
```
It looks normally distributed with few outliers above 100 

```{r echo=FALSE}
qplot(x = total.sulfur.dioxide, data = wine)
```
The total sulfur dioxide is normally distributed 

```{r echo=FALSE}
qplot(x = density, binwidth = 5e-04, data = wine)
```
Density is normally distributed with very few outliers

```{r echo=FALSE}
plot.hist('pH', wine, 0.01,'pH',"          PH Distribution")
```
The mean of the normal distribution is close to 3.2 


```{r echo=FALSE}
qplot(x = alcohol, data = wine)
```
```{r echo=FALSE}
qplot(x = alcohol, binwidth = 0.05, data = wine)
```
```{r echo=FALSE}
ggplot(aes(x = as.factor(quality)), data = wine)+
       geom_bar()+
       xlab("Wine Quality") + ylab("Count") +
       ggtitle("Bar Plot of Wine Quality")

```
The bar chart shows that the quality of the wine is good in most cases 

# Univariate Analysis

### What is the structure of your dataset?
This dataset contains information about wine. It includes 4898 observations of 12 variables.

### What is/are the main feature(s) of interest in your dataset?
Qualtiy is the main feature that we're intrested in.

### What other features in the dataset do you think will help support your \investigation into your feature(s) of interest?
There are some other interested features such as residual sugar, density, and volatile acidity.

### Did you create any new variables from existing variables in the dataset?
No

### Of the features you investigated, were there any unusual distributions? \Did you perform any operations on the data to tidy, adjust, or change the form \of the data? If so, why did you do this?
We took log 10 for the residual sugar since it was highly right skewed.

# Bivariate Plots Section

###Quality and Alcohol Density

```{r echo=FALSE, Bivariate_Plots}
qplot(alcohol, data = wine, binwidth = 0.2, fill = quality)

```
```{r echo=FALSE}
ggplot(wine, aes(factor(quality), density)) +
geom_boxplot()

```

```{r echo=FALSE}
ggplot(data= wine, aes(x = wine$residual.sugar, y = wine$density)) + 
  geom_point(alpha=0.1)
```
It's obvious that a positive correlation between residual sugar and density.
```{r echo=FALSE}
qplot(pH, data = wine, fill = sulphates)
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \investigation. How did the feature(s) of interest vary with other features in \the dataset?
we saw that alcohol and density are negatively correlated, meanwhile the relationship between density and residual sugar is positive.

### Did you observe any interesting relationships between the other features \(not the main feature(s) of interest)?
The relationship between PH and sulphates is weak


### What was the strongest relationship you found?
It was between density and residual sugar


# Multivariate Plots Section
```{r echo=FALSE}
ggplot(data= wine, aes(x = wine$density, y = wine$alcohol, color = as.factor(quality))) + 
geom_point() + scale_color_brewer(type='seq')

```
The scatter plot  shows a negative relationship between density and alcohol 

```{r echo=FALSE, Multivariate_Plots}
ggplot(data= wine, aes(x = wine$residual.sugar, y = wine$density, color = as.factor(quality))) + 
geom_point() + scale_color_brewer(type='seq')
```
The scatter plot shows a positive relationship between density and sugar

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \investigation. Were there features that strengthened each other in terms of \looking at your feature(s) of interest?
There is a strong relationship between quality and alcohol and density.

### Were there any interesting or surprising interactions between features?
surprisingly the residual sugar has a positive relationship.

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x = as.factor(quality)), data = wine)+
       geom_bar()+
       xlab("Wine Quality") + ylab("Count") +
       ggtitle("                                           Wine Quality")
```
### Description One
The distribution of the quality of white wine is normally distributed,also the quality of the wine lies between 5 and 7. 


### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(data= wine, aes(x = residual.sugar, y = density, color = as.factor(wine$quality))) +
geom_point() + scale_color_brewer(type='seq') +
labs(title="                    Density & Residual Sugar") +
xlab("Residual Sugar(g/dm^3) ") + 
ylab("Density (g/dm^3)")
```

### Description Two
The scatter plot shows that sugar and density have a positive correlation. 


### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(wine, aes(factor(quality), density)) +
geom_boxplot()+
  labs(title="                                    Density & Quality") +
  xlab('Qualtiy') +
  ylab('Density')
```

### Description Three
Density has strong negative correlation with the quality of the alcohol.

# Reflection
The dataset for the white wine contains information about wine features. It includes 4898 observations of 12 variables. The analysis was based on the relationships between variables, and after we analyzed some relationships we can say that good wine tends to have more alcohol and lower density.This analysis was a good experience for me.  
